Scale Space Edge Detection

### 2018-03-10 paper.tex
Started citations and added bib file

%&pdflatex
\documentclass{article}
\usepackage[final]{nips_2017}
\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{amsmath}
\usepackage{mathtools}
\usepackage{graphicx}
\usepackage{tabu}
\usepackage{mathrsfs}
\usepackage{morefloats}
\usepackage{float}
\usepackage{listings}
\usepackage[usenames,dvipsnames]{xcolor}
\usepackage{commath}
\usepackage{graphicx}
\usepackage{tikz,tikz-3dplot}
\usetikzlibrary{calc}
\usetikzlibrary{intersections}

@@

% Set the lstlistings to the Julia color scheme
\lstset{
  basicstyle       = \ttfamily,
  columns          = fullflexible,
  frame            = single,
  breaklines       = true,
  language         = Julia,
  basicstyle       = \ttfamily,
  keywordstyle     = \bfseries\color{blue},
  stringstyle      = \color{magenta},
  commentstyle     = \color{ForestGreen},
  showstringspaces = false,
  numbers          = left,
  xleftmargin      = 2em,
  framexleftmargin = 2em
}

@@

\section{Introduction}
Detecting edges in images has long been a problem with many unique solutions to approach it.
Generally speaking, the majority of algorithms are usually checking the image for some of the following features:
\begin{itemize}
\item large discontinuities in luminance values
\item discontinuities in different object orientations
\item large discontinuities in the intensity gradient
\end{itemize}
Common methods for detecting edges in images usually consist of convolving a differential matrix with the image that calculates the image's derivatives in the horizontal or vertical direction.
Algorithms like the Sobel, Canny, Prewitt, and Robert operators all work in a similar fashion, the only difference being which matrix they convolve with the image.
The problem with these methods is that they are not very adaptable. 
They generally will convolve a static matrix that is not dependent on the image, and could be potentially missing out on a lot of information. 
For example, given an image \textbf{F}, the following computes the gradient of the image in the $x$ and $y$ direction using the Sobel operator
\begin{equation}
	\textbf{G}_x =
	\begin{bmatrix}
		+1 &0 &-1 \\
		+2 &0 &-2 \\
		+1 &0 &-1 
	\end{bmatrix}
	* \textbf{F}, \;
	\textbf{G}_y =
	\begin{bmatrix}
		+1 &+2 &+1 \\
		0 &0 &0 \\
		-1 &-2 &-1 
	\end{bmatrix}
	* \textbf{F}
\end{equation}
This is why for our edge detection method, we will be using the scale space approach.
The benefit of using a scale space approach for edge detection, is we have the ability to classify the strength of the edges in the image.
This allows for a range of edges from very large immediate changes in intensity to very gradual.

\section{Motivation}

\section{Scale Space and Its Derivatives}
To understand how exactly we detect images in the scale space, we must first define what the scale space is.
If we have a continuous function of multiple variables such as $f(x,y)$, then we define the scale space representation of such a function as 
\begin{equation}
  L(x;t) = g(x,y;t) * f(x,y)
\end{equation}
Here $t$ represents the scale parameter, and can be thought of how much smoothing is applied to the function. The function $g$ is the Gaussian kernel given by
\begin{equation}
  g(x,y;t) = \frac{1}{2 \pi t}e^{-(x^2+y^2)/(2t)}
\end{equation}
With the scale space representation defined, we can now take derivatives of it as it is a continuous well-defined function.
Spatial derivatives are relatively simple being defined as the following
\begin{equation}
  L_{x^{\alpha}y^{\beta}}(\cdot;t) = \partial_{x^{\alpha}y^{\beta}}L(\cdot;t) = g_{x^{\alpha}y^{\beta}}(\cdot;t) * f(\cdot)
\end{equation}
However, when taking the partial derivative with respect to the scale $t$, it becomes more interesting.
The scale space representation collection is a solution for the diffusion equation. Therefore is has the useful property of 
\begin{equation}
  \partial_t L = \frac{1}{2} \nabla^2 L = \frac{1}{2} (\partial_{xx} + \partial_{yy})L
\end{equation}
with the initial condition of $L(x,y;0) = f(x,y)$.
So now, scale derivatives can be represented as spatial derivatives.

@@

The scale space representation is still defined in a similar fashion.
The following is the discrete version of the scale space operating on the function $f(x)$, which only has a single spatial variable.
\begin{equation}
  L(x;t) = (T(\cdot;t) * f(\cdot))(x;t)
\end{equation}
In this expression, $T$ represents the discrete version of the Gaussian kernel and is further evaluated as
\begin{equation}
  T(n;t) = e^{-t}I_n(t)
\end{equation}
where $I_n$ is the modified Bessel functions of integer order given by
\begin{equation}
  I_n(x) = i^{-n}J_{n}(ix) = \sum_{m=0}^{\infty}\frac{1}{m!\Gamma(m+n+1)}\left(\frac{x}{2}\right)^{2m+n}
\end{equation}
Now that the one dimensional case is understood for the scale space, we can expand this to two dimensions.
After all, images are compsed of two dimesnisons, so it makes sense that these operators can act on two dimensional functions.
The two dimensional scale space representation for discrete variables is given by the following
\begin{equation}
  L(x,y;t) = \sum_{m=-\infty}^{\infty}\sum_{n=-\infty}^{\infty}T(m;t)T(n;t)f(x-m,y-n)
\end{equation}
Even with the discrete case, the scale space representation must still satisfy the semidiscretized version of the diffusion equation.
Therefore by taking a scale derivative of the function we must have the following
\begin{equation}
  \partial_t L = \frac{1}{2}((1-\gamma)\nabla^2_5L+\gamma\nabla^2_\times L)
\end{equation}
where $\gamma \in [0,1]$ is a hyperparameter and,
\begin{equation}
  (\nabla^2_5f)_{0,0} = f_{-1,0} + f_{+1,0} + f_{0,-1} + f_{0,+1} - 4f_{0,0}
\end{equation}
\begin{equation}
  (\nabla^2_\times f)_{0,0} = \frac{1}{2}(f_{-1,-1} + f_{-1,+1} + f_{+1,-1} + f_{+1,+1} - 4f_{0,0})
\end{equation}
Note that $f_{-1,1}$ represents $f(x-1, y+1)$. Now that we have defined what discrete scale spaces and their derivatives look like, we can start to define what an edge is.

@@

\begin{table}[H]

@@

@@

@@

@@

@@

@@

@@

Here, the $v$-axis is parallel to the gradient direction at the point $(x_0, y_0)$, and the $u$-axis is perpendicular to the gradient direction at the same point. So then we define necessary angles as

@@

With our angles defined, we can then define our $(u,v)$ coordinate system by taking the respective partial derivatives of our spatial coordinate system.

@@

So then we can define our partials of $L$ with respect to this new $(u,v)$ coordinate system. So then we define an edge in this coordinate system with the following conditions.

@@

The way we can then represent these conditions in terms of spatial derivatives is as follows

@@

@@

Thus this adds two more conditions that must be satisified for an edge to be classified in the scale space. 

@@

The approach we took was to let our edge strength meteric be defined by the gradient magnitude that has been normalized by our $\gamma$ factor.
In this case we can define it as 
\begin{align}
  G_{\gamma}L &= L_{g,\gamma}^2 \\
  &= t^{\gamma}(L_x^2+L_y^2)
\end{align}
The first scale derivative of the gradient magnitude is calculated as follows
\begin{equation}
  \partial_t (G_{\gamma}L) = \gamma t^{\gamma-1}(L_x^2+L_y^2) + t^{\gamma}(L_x(L_{xxx}+L_{yyy})+L_y(L_{xxy} + L_{yyy}))
\end{equation}
Then the second scale derivative of the gradient magnitude is
\begin{equation}
  \begin{aligned}
    \partial_{tt} &= \gamma(\gamma-1)t^{\gamma-2}(L_x^2+L_y^2) \\
     &+ 2\gamma t^{\gamma-1}(L_x(L_{xyy}+L_{xxx})+L_y(L_{xxy}+L_{yyy})) \\
     &+\frac{t^{\gamma}}{2}\big((L_{xxx}+L_{xyy})^2+(L_{xxy}+L_{yyy})^2\big) \\
     &+ L_x(L_{xxxxx}+2L_{xxxyy}+L_{xyyyy}) \\
     &+L_y(L_{xxxxy}+2L_{xxyyy}+L_{yyyyy})\big)
  \end{aligned}
\end{equation}
Then the strength of the edges is found by computing the path integral over the maximal connected edge $\Gamma$ given by
\begin{equation}
  G(\Gamma) = \int_{(x;t) \in \Gamma} \sqrt{(G_{\gamma}L)(x;t)} \,\, \dif s
\end{equation}
where $\dif s^2=\dif x^2+\dif y^2$.

@@

\subsection{Marching Cubes}

We note that the equations $\partial_t (\mathcal G_{\text{$\gamma$-norm}} L) =0 $ and $L_{vv} = 0$ define isosurfaces $\mathcal Z_1$ and $\mathcal Z_2$, embedded in $L$.
Computing the intersection of these two surfaces and requiring the local maxima conditions $\partial_{tt} (\mathcal G_{\text{$\gamma$-norm}} L) < 0$ and $L_{vvv} < 0$ yields a family of isocurves, also embedded in $L$, which directly correspond to the edges in the image.

@@

%   \node[label=above:$P$,fill,circle,inner sep=1.75pt] at (P) {};
%   \node[label=below left:$Px$,fill,circle,inner sep=1.75pt] at (Px) {};
%   \node[label=below right:$Py$,fill,circle,inner sep=1.75pt] at (Py) {};
%   \node[label=above:$Pz$,fill,circle,inner sep=1.75pt] at (Pz) {};
%   \node[label=below:$Pxy$,fill,circle,inner sep=1.75pt] at (Pxy) {};
%   \node[label=above right:$Pyz$,fill,circle,inner sep=1.75pt] at (Pyz) {};
%   \node[label=above left:$Pxz$,fill,circle,inner sep=1.75pt] at (Pxz) {};
%   \node[label=above right:$O$,fill,circle,inner sep=1.75pt] at (O) {};
% \end{tikzpicture}
\section*{Appendix}

@@

\section*{References}

>>>>

%&pdflatex
\documentclass{article}
\usepackage[final]{nips_2017}
\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{amsmath}
\usepackage{mathtools}
\usepackage[draft]{graphicx}
\usepackage{tabu}
\usepackage{mathrsfs}
\usepackage{morefloats}
\usepackage{float}
\usepackage{listings}
\usepackage[usenames,dvipsnames]{xcolor}
\usepackage{commath}
\usepackage{graphicx}
\usepackage{tikz,tikz-3dplot}
\usepackage{longtable}
\usetikzlibrary{calc}
\usetikzlibrary{intersections}

@@

% Set the lstlistings to the Julia color scheme
\lstset{
  basicstyle       = \ttfamily,
  columns          = fullflexible,
  frame            = none,
  breaklines       = true,
  language         = Julia,
  basicstyle       = \ttfamily,
  keywordstyle     = \bfseries\color{blue},
  stringstyle      = \color{magenta},
  commentstyle     = \color{ForestGreen},
  showstringspaces = false,
  numbers          = left,
  xleftmargin      = 2em,
  framexleftmargin = 2em
}

@@

\section{Introduction}

Detecting edges in images has been a problem in image processing for a long time.
Generally speaking, the majority of algorithms are usually checking the image for some of the following features:
\begin{itemize}
\item large discontinuities in luminance values
\item discontinuities in different object orientations
\item large discontinuities in the intensity gradient
\end{itemize}
In this paper, we will be using the scale space representation of an image to approach the edge detection problem.
When we say scale, we refer to how much smoothing has been applied to the image. 
The theory behind the scale space is that detecting one-dimensional image features, like edges, may be dependent on what scale we operate at \cite{Lindeberg1998}. 
It follows that we can then define a scale space edge as a set of connected points that:
\begin{itemize}
\item have a gradient that takes on a local maximum value in the direction of the gradient\cite{Lindeberg1998}
\item have an edge strength metric which is a local maximum over various scales\cite{Lindeberg1998}.
\end{itemize}

\section{Motivation}
The motivation for this paper comes from observing the difference in quality of the scale space edge detection versus more conventional algorithms.
Common methods for detecting edges in images usually consist of convolving a discrete derivative matrix with the image that calculates the image's derivatives in the horizontal and vertical directions.
Algorithms like the Sobel, Canny, Roberts, and Prewitt do just exactly that. 
The problem with these methods is that they are not very adaptable. 
For example, given an image ${F}$, the following computes the gradient of the image in the $x$ and $y$ direction using the Sobel operator
\begin{equation}
	{G}_x =
	\begin{bmatrix}
		+1 &0 &-1 \\
		+2 &0 &-2 \\
		+1 &0 &-1 
	\end{bmatrix}
	* {F}, \;
	{G}_y =
	\begin{bmatrix}
		+1 &+2 &+1 \\
		0 &0 &0 \\
		-1 &-2 &-1 
	\end{bmatrix}
	* {F}
\end{equation}
Our algorithm will consist of convolving the ando5 matrix, which is more rotationally symmetric than the other operators listed above, with the image. 
This convolution will be be performed at multiple scales to evaluate the criteria necessary for classifying a point as an edge point in the scale space. 
We then run a marching cubes algorithm to connect the local maxima and list our edges based on strength. 
The ability to list edges and classify their strengths is the main advantage that the scale space edge algorithm has over conventional methods.

\section{Scale Space and Its Derivatives}
To understand how exactly we detect images in the scale space, we must first define what the scale space is.
If we have a continuous function of multiple variables such as $f(x,y)$, then we define the scale space representation of such a function as 
\begin{equation}
  L(x,y;t) = g(x,y;t) * f(x,y)
\end{equation}
Here $t$ represents the scale parameter, and can be thought of how much smoothing is applied to the function. The function $g$ is the Gaussian kernel given by
\begin{equation}
  g(x,y;t) = \frac{1}{2 \pi t}e^{-(x^2+y^2)/(2t)}
\end{equation}
With the scale space representation defined, we can now take derivatives of it as it is a continuous well-defined function.
Spatial derivatives are relatively simple being defined as the following
\begin{equation}
  L_{x^{\alpha}y^{\beta}}(\cdot;t) = \partial_{x^{\alpha}y^{\beta}}L(\cdot;t) = g_{x^{\alpha}y^{\beta}}(\cdot;t) * f(\cdot)
\end{equation}
However, when taking the partial derivative with respect to the scale $t$, it becomes more interesting.
The scale space representation collection is a solution for the diffusion equation. Therefore is has the useful property of \cite{Lindeberg1993}
\begin{equation}
  \partial_t L = \frac{1}{2} \nabla^2 L = \frac{1}{2} (\partial_{xx} + \partial_{yy})L
\end{equation}
with the initial condition of $L(x,y;0) = f(x,y)$.
So now, scale derivatives can be represented as spatial derivatives.

@@

The scale space representation is still defined in a similar fashion.
The following is the discrete version of the scale space operating on the function $f(x)$, which only has a single spatial variable.
\begin{equation}
  L(x;t) = (T(\cdot;t) * f(\cdot))(x;t)
\end{equation}
In this expression, $T$ represents the discrete version of the Gaussian kernel and is further evaluated as \cite{Lindeberg1993}
\begin{equation}
  T(n;t) = e^{-t}I_n(t)
\end{equation}
where $I_n$ is the modified Bessel functions of integer order given by
\begin{equation}
  I_n(x) = i^{-n}J_{n}(ix) = \sum_{m=0}^{\infty}\frac{1}{m!\Gamma(m+n+1)}\left(\frac{x}{2}\right)^{2m+n}
\end{equation}
Now that the one dimensional case is understood for the scale space, we can expand this to two dimensions.
After all, images are compsed of two dimesnisons, so it makes sense that these operators can act on two dimensional functions.
The two dimensional scale space representation for discrete variables is given by the following
\begin{equation}
  L(x,y;t) = \sum_{m=-\infty}^{\infty}\sum_{n=-\infty}^{\infty}T(m;t)T(n;t)f(x-m,y-n)
\end{equation}
Even with the discrete case, the scale space representation must still satisfy the semidiscretized version of the diffusion equation.
Therefore by taking a scale derivative of the function we must have the following \cite{Lindeberg1993}
\begin{equation}
  \partial_t L = \frac{1}{2}((1-\gamma)\nabla^2_5L+\gamma\nabla^2_\times L)
\end{equation}
where $\gamma \in [0,1]$ is a hyperparameter and,
\begin{equation}
  (\nabla^2_5f)_{0,0} = f_{-1,0} + f_{+1,0} + f_{0,-1} + f_{0,+1} - 4f_{0,0}
\end{equation}
\begin{equation}
  (\nabla^2_\times f)_{0,0} = \frac{1}{2}(f_{-1,-1} + f_{-1,+1} + f_{+1,-1} + f_{+1,+1} - 4f_{0,0})
\end{equation}
Note that $f_{-1,1}$ represents $f(x-1, y+1)$. Now that we have defined what discrete scale spaces and their derivatives look like, we can start to define what an edge is.

@@

\begin{table}[h]

@@

  \label{t_gma}

@@

  \label{t_tgma}

@@

  \label{t_ttgma}

@@

  \label{t_v}

@@

  \label{t_vv}

@@

  \label{t_vvv}

@@

Here, the $v$-axis is parallel to the gradient direction at the point $(x_0, y_0)$, and the $u$-axis is perpendicular to the gradient direction at the same point. So then we define necessary angles as \cite{Lindeberg1998}

@@

With our angles defined, we can then define our $(u,v)$ coordinate system by taking the respective partial derivatives of our spatial coordinate system

@@

So then we can define our partials of $L$ with respect to this new $(u,v)$ coordinate system. The results for $L_v$ can be found on page \pageref{t_v}. So then we define an edge in this coordinate system with the following conditions \cite{Lindeberg1998}

@@

The way we can then represent these conditions in terms of spatial derivatives is as follows \cite{Lindeberg1998}

@@

The results for $L_{vv}$ can be found on page \pageref{t_vv}.
The results for $L_{vvv}$ can be found on page \pageref{t_vvv}.

@@

Thus this adds two more conditions that must be satisified for an edge to be classified in the scale space \cite{Lindeberg1998}

@@

The approach we took was to let our edge strength metric be defined by the gradient magnitude that has been normalized by our $\gamma$ factor. 
Our results for the following calculations can be found on page \pageref{t_gma}.
In this case we can define our normalized gradient magnitude as \cite{Lindeberg1998}
\begin{align}
  G_{\gamma}L &= L_{v,\gamma}^2 \\
  &= t^{\gamma}(L_x^2+L_y^2)
\end{align}

The first scale derivative of the gradient magnitude results are found on page \pageref{t_tgma}. 
Its calculations are as follows \cite{Lindeberg1998}
\begin{equation}
  \partial_t (G_{\gamma}L) = \gamma t^{\gamma-1}(L_x^2+L_y^2) + t^{\gamma}(L_x(L_{xxx}+L_{yyy})+L_y(L_{xxy} + L_{yyy}))
\end{equation}
Then the second scale derivative of the gradient magnitude results are found on page \pageref{t_ttgma}.
Its calculations are as follows \cite{Lindeberg1998}
\begin{equation}
  \begin{aligned}
    \partial_{tt}(G_{\gamma}L) &= \gamma(\gamma-1)t^{\gamma-2}(L_x^2+L_y^2) \\
     & \hspace{.5pt}+ 2\gamma t^{\gamma-1}(L_x(L_{xyy}+L_{xxx})+L_y(L_{xxy}+L_{yyy})) \\
     & \hspace{.5pt}+\frac{t^{\gamma}}{2}\big((L_{xxx}+L_{xyy})^2+(L_{xxy}+L_{yyy})^2\big) \\
     & \hspace{.5pt}+ L_x(L_{xxxxx}+2L_{xxxyy}+L_{xyyyy}) \\
     & \hspace{.5pt}+ L_y(L_{xxxxy}+2L_{xxyyy}+L_{yyyyy})\big)
  \end{aligned}
\end{equation}
Then the strength of the edges is found by computing the path integral over the maximal connected edge $\Gamma$ given by \cite{Lindeberg1998}
\begin{equation}
  G(\Gamma) = \int_{(x;t) \in \Gamma} \sqrt{(G_{\gamma}L)(x;t)} \,\, \dif s
\end{equation}
where $\dif s^2=\dif x^2+\dif y^2$.

We then perform a marching cubes algorithm that will connect local points that we have identified as edge points. The details for this can be found on page \pageref{mcube}.

@@

\subsection{Marching Cubes}
\label{mcube}
We note that the equations $\partial_t (\mathcal G_{\text{$\gamma$-norm}} L) =0 $ and $L_{vv} = 0$ define isosurfaces $\mathcal Z_1$ and $\mathcal Z_2$, embedded in $L$.
Computing the intersection of these two surfaces and requiring the local maxima conditions $\partial_{tt} (\mathcal G_{\text{$\gamma$-norm}} L) < 0$ and $L_{vvv} < 0$ yields a family of isocurves, also embedded in $L$, which directly correspond to the edges in the image.

@@

%   \node[label=above:$P$,fill,circle,inner sep=1.75pt] at (P) {};
%   \node[label=below left:$Px$,fill,circle,inner sep=1.75pt] at (Px) {};
%   \node[label=below right:$Py$,fill,circle,inner sep=1.75pt] at (Py) {};
%   \node[label=above:$Pz$,fill,circle,inner sep=1.75pt] at (Pz) {};
%   \node[label=below:$Pxy$,fill,circle,inner sep=1.75pt] at (Pxy) {};
%   \node[label=above right:$Pyz$,fill,circle,inner sep=1.75pt] at (Pyz) {};
%   \node[label=above left:$Pxz$,fill,circle,inner sep=1.75pt] at (Pxz) {};
%   \node[label=above right:$O$,fill,circle,inner sep=1.75pt] at (O) {};
% \end{tikzpicture}

\section*{Appendix}

@@

\bibliography{bibliography}{}
\bibliographystyle{plain}
